### 2019-04-29 papers/haskell19/6-casestudies.tex
WIP case studies

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{tikzpicture}
  \begin{axis}
    [ ybar
    , enlargelimits=0.15
    , enlarge x limits=0.3
    , /pgf/bar width=15pt
    , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
    , ylabel={Relative Generation Time}
    , symbolic x coords={SExp, RBT, Html}
    , xtick=data
    % , nodes near coords
    , nodes near coords align={vertical}
    ]
    \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
    \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
    \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
    \legend{Manual, Modular/Linear, Modular/Balanced}
\end{axis}
\end{tikzpicture}

>>>>

The random generation framework presented throughout this work allows us to
write extensible generators in a very concise way.
%
However, this expresiveness comes attached to a perceptible runtime overhead.
%
In this section we evaluate the implicit cost of composing generators using
three real-world case studies.

As we have shown in Section \ref{sec:generators}, the random generation
process we propose in this paper can be seen as having two phases.
%
First, generate random values from the representation types used to specify the
shape of our data; and then we use their algebras to translate them to the
corresponding values of our target data types.
%
In particular, the tranformation step is expected to pattern match repeatedly
against the |InL| and |InR| constructors of the |oplus| operators when
traversing each construction injection.
%
Because of this, in general we expect a performance impact with respect to
manually-written concrete generators.

As recently analyzed by \citeauthor{KiriyamaOptimizingDTC}, this slowdown is
expected to be linear in the depth of our representation type
\cite{KiriyamaOptimizingDTC}.
%
Hence, one can drastically reduce the runtime overhead by associating each
|oplus| operator in a balanced fashion.
%
So, for instance, instead of writing |(f oplus'' g oplus'' h oplus'' i)|, which
is implicitly parsed as |(f oplus'' (g oplus'' (h oplus'' i)))|; we can
associate constructions as |((f oplus'' g) oplus'' (h oplus'' i))|, thus
reducing the depth of our representation from four to three levels and, in
general, from a $\mathcal{O}(n)$ to a $\mathcal{O}(log(n))$ complexity in the
runtime overhead, where $n$ is the amount of constructions under consideration.
%
To apply this optimization automatically, and relying on more sophisticated
techniques like prefix-free encodings to obtain optimal representations, is a
undergoing path of our future work.

To evaluate our approach, we analyzed the performance of generating random
values using three real-world case studies:
\begin{inparaenum}[(i)]
\item Red-Black Trees, inspired by Okasaki's formulation,
\item Lisp S-expressions, inspired by the package \emph{hs-zuramaru}, and
\item HTML expressions, inspired by the \emph{html} package, and following the
  same structure as our motivating |Html| example.
\end{inparaenum}
%
The expanse of each case study can be summarized as follows:

\begin{table}[H]
  \begin{tabular}{||c||c||c||c||c||}
    \hline
    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
    \hline
    RBT        & 2      & 6      & 5      & 13                  \\ \hline
    SExp       & 6      & 9      & -      & 15                  \\ \hline
    Html       & 4      & -      & 132    & 136                 \\ \hline
  \end{tabular}
\end{table}

For each case study, we benchmarked the execution time required to generate 1000
random values using both manually-written concrete generators, and modular
generators obtained using our approach.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,5}
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, Html}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
      \legend{Manual, Modular/Linear, Modular/Balanced}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}
